A Statistical Dynamical Model to Predict Extreme Events and Anomalous Features in Shallow Water Waves with Abrupt Depth Change

### 2018-11-27 paper1_tKdV/tKdV_theory.tex
NM

The surface wave turbulence is modeled by a one-dimensional deterministic
dynamical model. 
\todo[inline]{This first sentence seems awkward to me.}
The Korteweg-de Vries (KdV) equation \cite{johnson1997modern}
is a leading-order approximation of the surface waves that are determined
by the balance of nonlinear and dispersive effects in an appropriate
far-field limit. Here, the KdV equation is truncated in the first
$\Lambda$ modes (with $J=2\Lambda+1$ grid points) to generate weakly
turbulent dynamics. The surface displacement is described by
the state variable $u_{\Lambda}^{\pm}\left(x,t\right)$ with superscript `$-$'
for the incoming waves and `$+$' for the outgoing waves. The Galerkin
truncated variable $u_{\Lambda}=\sum_{1\leq\left|k\right|\leq\Lambda}\hat{u}_{k}\left(t\right)e^{ikx}$
is normalized with zero mean $\hat{u}_{0}=0$ and unit energy $2\pi\sum_{k=1}^{\Lambda}\left|\hat{u}_{k}\right|^{2}=1$,
which are conserved quantities, and $u_{\Lambda}\equiv\mathcal{P}_{\Lambda}u$
denotes the subspace projection. The evolution of  $u_{\Lambda}^{\pm}$ is governed by the truncated
KdV equation with depth change $D_{\pm}$
\begin{equation}
\frac{\partial u_{\Lambda}^{\pm}}{\partial t}+\frac{D_{\pm}^{-3/2}}{2}E_{0}^{1/2}L_{0}^{-3/2}\frac{\partial}{\partial x}\mathcal{P}_{\Lambda}\left(u_{\Lambda}^{\pm}\right)^{2}+D_{\pm}^{1/2}L_{0}^{-3}\frac{\partial^{3}u_{\Lambda}^{\pm}}{\partial x^{3}}=0,\label{eq:tKdV_nondim}
\end{equation}
on the normalized periodic domain $x\in\left[-\pi,\pi\right]$. The conserved Hamiltonian can be decomposed into a cubic and a quadratic term
\begin{align*}
\mathcal{H}_{\Lambda}^{\pm} & =  D_{\pm}^{-3/2}E_{0}^{1/2}L_{0}^{-3/2}H_{3}\left(u_{\Lambda}^{\pm}\right)-D_{\pm}^{1/2}L_{0}^{-3}H_{2}\left(u_{\Lambda}^{\pm}\right),\\
& H_{3}\left(u\right) =  \frac{1}{6}\int_{-\pi}^{\pi}u^{3}dx,\;H_{2}\left(u\right)=\frac{1}{2}\int_{-\pi}^{\pi}\left(\frac{\partial u}{\partial x}\right)^{2}dx.
\end{align*}
Equation [\ref{eq:tKdV_nondim}] is non-dimensionalized in the periodic
domain. The depth is assumed to be unit $D_{-}=1$
before the ADC and becomes $D_{+}<1$
after the ADC. We introduce parameters $\left(E_{0},L_{0},\Lambda\right)$
based on the following assumptions:
\begin{itemize}
\item The wavenumber truncation $\Lambda$ is fixed in a moderate value
for generating weakly turbulent dynamics;
\item The state variable $u_{\Lambda}^{\pm}$ is normalized with zero mean
and unit energy, $\mathcal{M}\left(u_{\Lambda}\right)=0,\mathcal{E}\left(u_{\Lambda}\right)=1$,
which are conserved during evolution. Meanwhile, $E_{0}$ characterizes the total
energy injected into the system based on the paddle amplitude $\left(\Delta\theta\right)^{2}$;
\item The length scale of the system is denoted $L_{0}$. The value is
chosen so that the resolved scale $\Delta x=2\pi L_{0}/J$ is comparable
to the the characteristic wave length $\lambda_{c}$ in
the experiments.
\end{itemize}
% ORIGINAL
%The intuition for the distinct model dynamics comes from the balance
%between the cubic and quadratic terms in the Hamiltonian $\mathcal{H}_{\Lambda}^{\pm}$.
%After the depth change, $D_{+}<1$, more weight is added in the cubic
%term, $H_{3}$, for stronger nonlinearity and weaker dispersion for
%the third-order derivative term reflected by the smaller coefficient
%for $H_{2}$ in the Hamiltonian. 
Some intuition for how \eqref{eq:tKdV_nondim} produces different dynamics on either side of the ADC can be gained by considering the balance of the cubic and quadratic terms in the Hamiltonian $\mathcal{H}_{\Lambda}^{\pm}$. The depth change, $D_{+}<1$, increases the weight of $H_3$ and decreases that of $H_2$, thus promoting the effects of nonlinearity and reducing dispersion.
Since $\frac{\partial u}{\partial x}$ is the slope of the wave height, $H_{2}\left(u\right)$ measures the
wave slope energy.

A \emph{deterministic matching condition} is applied to the surface displacement $u_{\Lambda}^{\pm}$ to link the incoming and outgoing wave trains. Assuming the abrupt depth change is met at $t=T_{\mathrm{ADC}}$, the matching condition is given by
\[
u_{\Lambda}^{-}\left(x,t\right)\mid_{t=T_{\mathrm{ADC}}-}=u_{\Lambda}^{+}\left(x,t\right)\mid_{t=T_{\mathrm{ADC}}+},
\]
Equation [\ref{eq:tKdV_nondim}] is not designed to capture the short
scale changes in rapid time. 
Rather, since we are interested in modeling statistics before and after the depth change, we will examine the long-time dynamics of large-scale structures.

\todo[inline]{I modified the last sentence above. Feel free to revert to your original sentence if you do not like mine. (the change looks good)}

>>>>

The Korteweg-de Vries (KdV) equation \cite{johnson1997modern}
is a leading-order approximation of surface waves that are determined
by the balance of nonlinear and dispersive effects in an appropriate
far-field limit. Here, the KdV equation is truncated in the first
$\Lambda$ modes (with $J=2\Lambda+1$ grid points) to generate weakly
turbulent dynamics. The surface displacement is described by
the state variable $u_{\Lambda}^{\pm}\left(x,t\right)$ with superscript `$-$'
for the incoming waves and `$+$' for the outgoing waves. The Galerkin
truncated variable $u_{\Lambda}=\sum_{1\leq\left|k\right|\leq\Lambda}\hat{u}_{k}\left(t\right)e^{ikx}$
is normalized with zero mean $\hat{u}_{0}=0$ and unit energy $2\pi\sum_{k=1}^{\Lambda}\left|\hat{u}_{k}\right|^{2}=1$,
which are conserved quantities, and $u_{\Lambda}\equiv\mathcal{P}_{\Lambda}u$
denotes the subspace projection. The evolution of  $u_{\Lambda}^{\pm}$ is governed by the truncated
KdV equation with depth change $D_{\pm}$
\begin{equation}
\frac{\partial u_{\Lambda}^{\pm}}{\partial t}+\frac{D_{\pm}^{-3/2}}{2}E_{0}^{1/2}L_{0}^{-3/2}\frac{\partial}{\partial x}\mathcal{P}_{\Lambda}\left(u_{\Lambda}^{\pm}\right)^{2}+D_{\pm}^{1/2}L_{0}^{-3}\frac{\partial^{3}u_{\Lambda}^{\pm}}{\partial x^{3}}=0. \label{eq:tKdV_nondim}
\end{equation}
Equation [\ref{eq:tKdV_nondim}] is non-dimensionalized on the periodic
domain $x\in\left[-\pi,\pi\right]$. The depth is assumed to be unit $D_{-}=1$ before the ADC and becomes $D_{+}<1$ after the ADC. 
The conserved Hamiltonian can be decomposed into a cubic and a quadratic term
\begin{align*}
\mathcal{H}_{\Lambda}^{\pm} & =  D_{\pm}^{-3/2}E_{0}^{1/2}L_{0}^{-3/2}H_{3}\left(u_{\Lambda}^{\pm}\right)-D_{\pm}^{1/2}L_{0}^{-3}H_{2}\left(u_{\Lambda}^{\pm}\right),\\
& H_{3}\left(u\right) =  \frac{1}{6}\int_{-\pi}^{\pi}u^{3}dx,\;H_{2}\left(u\right)=\frac{1}{2}\int_{-\pi}^{\pi}\left(\frac{\partial u}{\partial x}\right)^{2}dx.
\end{align*}
We introduce parameters $\left(E_{0},L_{0},\Lambda\right)$
based on the following assumptions:
\begin{itemize}
\item The wavenumber truncation $\Lambda$ is fixed at a moderate value
for generating weakly turbulent dynamics;
\item The state variable $u_{\Lambda}^{\pm}$ is normalized with zero mean
and unit energy, $\mathcal{M}\left(u_{\Lambda}\right)=0,\mathcal{E}\left(u_{\Lambda}\right)=1$,
which are conserved during evolution. Meanwhile, $E_{0}$ characterizes the total
energy injected into the system based on the driving amplitude $\Delta \theta$;
\item The length scale of the system $L_{0}$ is chosen so that the resolved scale $\Delta x=2\pi L_{0}/J$ is comparable to the the characteristic wave length $\lambda_{c}$ from the experiments.
\end{itemize}
% ORIGINAL
%The intuition for the distinct model dynamics comes from the balance
%between the cubic and quadratic terms in the Hamiltonian $\mathcal{H}_{\Lambda}^{\pm}$.
%After the depth change, $D_{+}<1$, more weight is added in the cubic
%term, $H_{3}$, for stronger nonlinearity and weaker dispersion for
%the third-order derivative term reflected by the smaller coefficient
%for $H_{2}$ in the Hamiltonian. 
Some intuition for how \eqref{eq:tKdV_nondim} produces different dynamics on either side of the ADC can be gained by considering the relative contributions of the cubic and quadratic terms to the Hamiltonian $\mathcal{H}_{\Lambda}^{\pm}$. The depth change, $D_{+}<1$, increases the weight of $H_3$ and decreases that of $H_2$, thus promoting the effects of nonlinearity and reducing dispersion.
Since $\frac{\partial u}{\partial x}$ is the slope of the wave height, $H_{2}\left(u\right)$ measures the
wave slope energy.

A \emph{deterministic matching condition} is applied to the surface displacement $u_{\Lambda}^{\pm}$ to link the incoming and outgoing wave trains. Assuming the abrupt depth change is met at $t=T_{\mathrm{ADC}}$, the matching condition is given by
\[
u_{\Lambda}^{-}\left(x,t\right)\mid_{t=T_{\mathrm{ADC}}-}=u_{\Lambda}^{+}\left(x,t\right)\mid_{t=T_{\mathrm{ADC}}+},
\]
Equation [\ref{eq:tKdV_nondim}] is not designed to capture the short
scale changes in rapid time. 
Rather, since we are interested in modeling statistics before and after the ADC, we will examine the long-time dynamics of large-scale structures.
